Distributed Tic Tac Toe:\\ Peer-to-Peer Team-Based Gaming System

### 2017-11-19 paper/paper.tex
Article Additions
"Article Guide" document to help fill out each section, as per the assignment spec.

"References" document to keep track of references along with citation number until we can index it ourselves.

Small changes to the paper to make it less informal. Still more needed on this.

Small additions to Introduction to argue "novel contribution" and high level overview of the game in the "System Design" section.

\section{Introduction}
Coding nowadays is the main occupation of loads of people and it is also a
past-time for just as many. Most people get their first steps in writing code
following the local single-threaded model. Often, it can be difficult to
beginner programmers or even those with more experience to abstract the key
knowledge they have obtained using this model and apply it in a different way.
Distributed software requires a new look on the system – the number of devices
has gone up and with it so has the number of processes and other concerns. Some
concepts have now changed and need to be dealt with differently – one such
example is mutual exclusion. We believe programs that serve as examples and
focus on simple local tasks turned into distributed ones can help many people
trying to make this step with a hands-on experience of what they are preparing
for. Seeing theory in practice is a proven technique and there is evidence for
this in every university course that features practical work. Therefore, we have
chosen to build a distributed game of tic-tac-toe – a game known to possibly
everyone. By restraining from introducing additional complexity as part of the
goal of the project, the focus is shifted towards the implications of the
distributive aspect of the system and the way they are handled. The finished
product also has another purpose – it is an actual game that can be played and
as such it can bring joy to people and increase their curiosity about computers.

@@

\section{System Design}

@@

\section{Related Work}
While other multiplayer distributed games already exist, they have not been
created with this project’s task in mind and therefore do not serve the same
purpose. They can be split into two categories – very unprofessional and too
professional. For example, hobby implementations exist that deal with
distributed problems in a bad fashion. Similarly, a single-player-per-team
distributed tic-tac-toe game in Java can be found online that does not comply
with any design guidelines and puts all functionality in a single class and as
such cannot teach students good practice and might confuse them further about
distributive programming \cite{2}. Separately, very in-depth materials that
provide state-of-the-art practices can be found \cite{1} but there is a high
possibility that they might be incomprehensible to a beginner reader looking to
tackle basic distributed concepts to begin with. Furthermore, none of the
similar implementations residing on the Internet take advantage of the Java RMI
API which must be considered a good starting point for distributed systems
learners for it to be taught as part of our degree and is a requirement for this
project. Instead, they use different sets of other technologies which can
further confuse their reader and introduce a steeper learning curve.

>>>>

\section{Introduction}
Coding is the main occupation of many people nowadays, and is also a
past-time for many too. Most people get their first steps in writing code
following the local single-threaded model. Often, it can be difficult for
beginner programmers, or even those with more experience, to abstract the key
knowledge they have obtained using this model and then apply it in a different way.

Distributed software requires a new look on the system – the number of devices
has gone up and with it so have the number of processes, concerns and complexities.
Some concepts have now changed and need to be dealt with differently; One such
example if this would be mutual exclusion. We believe programs that serve as examples and
focus on simple local tasks turned into distributed ones can help many people
trying to make this step with a hands-on experience of what they are preparing
for. Seeing theory in practice is a proven technique and there is evidence for
this in every university course that features practical work. This is especially
true where games are concerned, proving to be an effective means of communicating 
programming methods and techniques in an engaging way \cite{3}. For this reason, 
the team has decided to take a simple, well known game and develop it into a distributed 
system; Tic Tac Toe played as a peer-to-peer, team-based gaming system where the players are organised into teams and required to vote on moves for their team.

By refraining from introducing additional complexity as part of the goal of the project, and 
allowing the users to see the distributed nature of the software being demonstrated via a simple game, the focus is shifted towards the implications of the distributive aspects of the system and 
the way they are handled. The finished product can serve to entertain the end users, as well as increase their curiosity and insight about computers and distributed systems. Using Tic Tac Toe as a foundation to demonstrate a distributed system benefits from both the simplicity and familiarity of the game itself, meaning no complex rulesets or in-depth mechanisms need be explained to the users, allowing the distributed nature to be a novel way of experiencing a widely known game.

@@

\section{System Design}
Tic Tac Toe is a simple game played on a grid of 3x3, consisting of two players. One player uses an “X” marker and the other player uses an “O” marker to select a grid space to occupy. The objective of the game is to take turns until one player manages to occupy three grid spaces in a straight line, be it horizontal, vertical or diagonal.

The proposed design takes this simple concept and uses it to demonstrate a distributed system, whereby the two “players” are instead two teams of up to 5 players, which take turns to vote on where their team should place their markers. Each team has a designated "leader" which communicates data with their team's peers, and the opposing team's "leader" to synchronise current gameplay state. The leaders communicate the vote of their team to the opposing leader so that they may pass that result on to the peers of that team.

@@

\section{Related Work}
While other multiplayer distributed games already exist, they have not been
created with this project’s task in mind and therefore do not serve the same
purpose. They can be split into two categories – very unprofessional and too
professional. For example, hobby implementations exist that deal with
distributed problems in a bad fashion, promoting bad and incorrect programming 
practises. Similarly, a single-player-per-team distributed tic-tac-toe game in 
Java can be found online that does not comply with any design guidelines and 
puts all functionality in a single class and as such, cannot teach students good 
practice and has the potential of confusing the learner further about distributive 
programming \cite{2}.

Separately, very in-depth materials that provide state-of-the-art practices can 
be found \cite{1} but there is a high possibility that they might be incomprehensible 
to a beginner reader looking to tackle basic distributed concepts to begin with. 
Furthermore, none of the similar implementations residing on the Internet take advantage 
of the Java RMI API which must be considered a good starting point for distributed systems
learners for it to be taught as part of our degree and is a requirement for this
project. Instead, they use different sets of other technologies which can
further confuse their reader and introduce a steeper learning curve.
